Algebraic Principles for Concurrency Verification Tools

### 2013-10-07 Paper/paper.tex
CNPq
More on model. Owicki-Gries doesn't work

\documentclass{llncs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{graphicx,color}
\usepackage{multicol}
\usepackage{bussproofs}
\usepackage{float}

@@

\DeclareFontFamily{U}{wncy}{}
\DeclareFontShape{U}{wncy}{m}{n}{<->wncyss10}{}
\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
\DeclareMathSymbol{\sh}{\mathord}{mcy}{"78} 
\DeclareMathOperator{\sha}{\sh}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\fin}{finite}

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Kleene Algebra}
\label{sec:KA}

A \emph{dioid} or \emph{idempotent semiring} is an structure
$(D,+,\cdot,0,1)$ such that $(D,+,0)$ is a join semilattice with zero,
$(D,\cdot,1)$ is a monoid, $+$ distributes over $\cdot$ from both the
right and left, and $0$ is a right and left annihilator for
$\cdot$. The natural partial order $\le$ is defined as $x \le y $ iff
$x + y = y$.

A Kleene algebra extends a dioid with a star operation satisfying the
following laws.
\begin{align*}
  1 + xx^* &\le x^* &\qquad z + xy \le y &\implies x^*z \le y\\
  1 + x^*x &\le x^* &\qquad z + yx \le y &\implies zx^* \le y
\end{align*}

In the context of programs, one typically thinks of $\cdot$ as being
sequential composition, $+$ as non-deterministic choice, $0$ as the
abortive action, $1$ as skip and $1$ as finite iteration. For
concurrent software, it is natural to extend the notion of a dioid to
that of a \emph{trioid} with operators for both sequential and
parallel composition. One can also axiomatise a parallel star
operation, but this seems to be a less useful operation. A
dioid or trioid is \emph{weak} if it does not satisfy $x0 = 0$.

Formally, a trioid is a structure $(D,+,\cdot,\|,0,1)$, such that
both $(D,+,\cdot,0,1)$ and $(D,+,\|,0,1)$ are dioids. Previous work in
this area has focused on identifying laws linking the sequential and
parallel composition operators, notably, the exchange law
\begin{align*}
  (a \| b) \cdot (c \| d) \le (a \cdot d) \| (b \cdot c).
\end{align*}
The exchange law is appealing as it provides a way to derive many
properties of parallel composition from a single law. For example, one
can show that $\|$ must be commutative, that $x \cdot y \le x \| y$

@@

Define a rely-guarantee algebra as a structure
$(K,RG,+,\sqcap,\cdot,\|,^\star,0,1)$, where $(K,+,\sqcap)$ is a
lattice, $(K,+,\cdot,\|,0,1)$ is a weak trioid and
$(K,+,\cdot,^\star,0,1)$ is a Kleene algebra. $RG$ is a distinguished
subset of relys/guarantees which satisfy the following axioms
\begin{align}
r\|r &\le r, \label{rg1}\\
r &\le r\|r', \label{rg2}\\
r\|xy &= (r\|x)(r\|y), \label{rg3}\\
r\|x^+ &\le (r\|x)^+ \label{rg4}.
\end{align}
By convention, we use $r$ and $g$ to refer to elements of $RG$, and
$x,y,z$ for arbitrary elements of $K$. Some elementary consequences of these rules are as follows
\begin{align*}
1 &\le r,\\
r^\star = rr &= r = r\|r,\\
r\|x^+ &= (r\|x)^+.
\end{align*}

@@

Axiom (\ref{rg1}) states that interference from the a rely condition
being run twice in parallel, is no different from just the
interference from that rely condition. Axiom (\ref{rg2}) states that
interference from a single rely condition is less than interference
from itself and another rely condition. Axiom (\ref{rg3}) allows a
rely or guarantee to be split across sequential programs. Axiom
(\ref{rg4}) is similar to Axiom (\ref{rg3}) in intent, except it deals
with finite iteration.

@@

Axiom (\ref{rg4}) can be derived from (\ref{rg3}) in all finite
models. This is because any finite $K$ is a weak left quantale with
respect to both sequential and parallel composition. In a quantale,
fixpoint fusion and induction laws can be used to prove that
\begin{align*}
r\|x^+ = r\|(\mu y.\; x + xy) = \mu y.\; r\|x + r\|xy \le (r\|x)^+.
\end{align*}
Thus it is impossible to construct a finite model demonstrating
that (\ref{rg4}) is independent from (\ref{rg1}) -- (\ref{rg2}).

@@

Jones quintuples can be encodeded in this setting as
\begin{align}
r, g \vdash \{p\} c \{q\} \iff p(r\|c) \le q \land c \le g.
\end{align}
Using this encoding we can derive all the standard rules for
rely-guarantee, shown in Figure \ref{fig:rgrules}.

@@

\begin{prooftree}
\RightLabel{Star}
\AxiomC{$pr \le p$}
\AxiomC{$r, g \vdash \{p\}x\{p\}$}
\BinaryInfC{$r, g \vdash \{p\}x^\star\{p\}$}
\end{prooftree}
\caption{Rely-guarantee inference rules}
\label{fig:rgrules}
\end{figure}
x

x
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Concurrent Kleene Algebra with Tests}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Language Model}

In this section, we construct a finite language model satisfying the
axioms in Section \ref{sec:RG}. Restricting our attention to finite
languages means that we do not need to concern ourselves with
termination side-conditions.

As usual, the product of two languages X and Y over an alphabet
$\Sigma$ is defined as $XY = \{x\frown y.\; x \in X \land y \in Y\}$.
The language product is associative, and distributes over arbitrary
joins from both the left and the right. The shuffle $\sha$ on a pair
of finite words is defined inductively as
\begin{align*}
\epsilon \sha s &= \{s\},\\
s \sha \epsilon &= \{s\},\\
as \sha bt &= a(s \sha bt) \cup b(as \sha t),
\end{align*}
which is then lifted to languages X and Y as the shuffle product,
\begin{align*}
X \sha Y = \bigcup\{xs \sha ys.\; x \in X \land x \in Y\}.
\end{align*}
The shuffle product is associative, commutative, and distributes over
arbitrary joins. Both products share the same unit, $\{\epsilon\}$ and
zero, $\emptyset$. Therefore, $(\mathcal{P}(\Sigma)^*), \cup, \cdot,
\sha, \emptyset, \{\epsilon\})$ forms a weak trioid.

\begin{lem}
$(\mathcal{P}(\Sigma)^*), \cup, \cdot, \sha, \emptyset, \{\epsilon\})$ is a weak trioid.
\end{lem}

Our alphabet comprises relations ($R,S,T,\dots$) over a state space
$\Sigma$. An example of a word in
$(\{\sigma_1,\sigma_2,\sigma_3\}^2)^*$ is shown below. This word only
has a single valid \emph{trace}, which is
$\sigma_1\rightarrow\sigma_1\rightarrow\sigma_2\rightarrow\sigma_3\rightarrow\sigma_2$.
We define an operator $\tr(X)$ which returns all the traces in a language
$X$.

@@

Following~\cite{brookes_full_1993} and~\cite{dingel_refinement_2002},
we inductively generate the stutter/mumble language $w^\dagger$ for a
word $w$ in a language over $\Sigma^2$ as follows. Assume $R,S,T \in
\Sigma^2$ and $u,v,w \in (\Sigma^2)^*$. First, $Iw \in
w^\dagger$. Secondly, if $uRv \in x^\dagger$ then $uIRv \in
x^\dagger$ and $uRIv \in x^\dagger$ (\emph{stuttering}). Thirdly, if
$uRSv \in w^\dagger$ then $u(R;S)v \in w^\dagger$
(\emph{mumbling}). The stutter/mumble closure of a language is
defined by lifting this operation to languages in the obvious way.

Next, we define the following language for a word $w = R_0R_1\dots
R_n$. The language $w^\udagger$ contains all strings $S_0S_1\dots S_n$
such that for all $m \le n$, $S_m \subseteq R_m$. Again this operation
is lifted to languages in the same way as before. Both these closure
operators are combined then into a single closure operator,
$X^\ddagger$, which satisfies the properties
\begin{align*}
(X^\ddagger \sha Y)^\ddagger &= (X \sha Y)^\ddagger,\\
(X^\ddagger Y)^\ddagger = (&XY)^\ddagger = (XY^\ddagger)^\ddagger,\\
(X \cup Y)^\ddagger &= (X^\ddagger \cup Y^\ddagger),\\
(X^\ddagger \cap Y^\ddagger)^\ddagger &= (X^\ddagger \cap Y^\ddagger),\\
\tr(X^\ddagger) &\subseteq X^\ddagger.
\end{align*}
For all binary operators $\bullet$, we can therefore define
$\bullet_\ddagger$ as $\lambda X Y. (X^\ddagger\bullet
Y^\ddagger)^\ddagger$ and show that
$(\mathcal{P}((\Sigma^2)^*)^\ddagger,\cup_\ddagger,\cdot_\ddagger,\sha_\ddagger,\emptyset,{\epsilon}^\ddagger)$
is a weak trioid.

@@

>>>>

\documentclass{llncs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{graphicx,color}
\usepackage{multicol}
\usepackage{bussproofs}
\usepackage{float}
\usepackage{centernot}

@@

\DeclareFontFamily{U}{wncy}{}
\DeclareFontShape{U}{wncy}{m}{n}{<->wncyss10}{}
\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
\DeclareMathSymbol{\sh}{\mathord}{mcy}{"78} 
\DeclareMathOperator{\sha}{\sh}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\fin}{finite}
\DeclareMathOperator{\rely}{rely}
\DeclareMathOperator{\guar}{guar}

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Semirings and Kleene Algebra}
\label{sec:KA}

A \emph{semiring} is an structure
$(D,+,\cdot,0,1)$ satisfying the following laws,
\begin{align*}
  x(y + z) &= xy + xz, &\qquad (x + y)z &= xz + yz,\\
  (x + y) + z &= x + (y + z), &\qquad (xy)z &= x(yz),\\
  x + y &= y + x, &\qquad 0 + x &= x,\\
  1x &= x, &\qquad x1 &= x,\\
  0x &= 0, &\qquad x0 &= 0.
\end{align*}
A semiring is \emph{idempotent} if $x + x = x$, and \emph{weak} if $x0
= 0$ does not hold. The natural partial order $\le$ is defined as $x
\le y $ iff $x + y = y$. Idempotent semirings are also called
\emph{dioids}. In the context of programs, one typically thinks of
$\cdot$ as being sequential composition, $+$ as non-deterministic
choice, $0$ as the abortive action and $1$ as skip. For concurrent
software, it is natural to extend the notion of a dioid to that of a
\emph{trioid} with operators for both sequential and parallel
composition. Formally, a trioid is a structure $(D,+,\cdot,\|,0,1)$
where $(D,+,\cdot,0,1)$ is a dioid and,
\begin{align*}
  x\|(y + z) &= x\|y + x\|z, &\qquad x\|y &= y\|x,\\
  (x\|y)\|z &= x\|(y\|z), &\qquad x\|1 &= x,\\
  0\|x &= x.
\end{align*}

In a trioid there are no laws governing the interaction between
sequential and parallel composition. One law which has been considered
in previous papers is the \emph{exchange} law
\begin{align*}
  (w \| x) \cdot (y \| z) \le (w \cdot z) \| (x \cdot y).
\end{align*}

The exchange law is initially appealing as it provides a way to derive
many properties of parallel composition from a single law. For example
one can show that $\|$ must be commutative, or that $x\cdot y \le
x\|y$. However, there are many interesting models which do not satisfy
this law. First, one could consider fair parallel composition
$x\parallel_f y$ in a model with infinite, or non-terminating
programs. In this model, $x \cdot y \not\leq x \parallel_f y$, in the
case where $x$ is non-terminating. A second example would be
predicate-transformer style models for separation logic, where $\|$ is
interpreted as separating conjunction. In such models, $\|$ and
$\cdot$ have different units, which, in the presence of the exchange
law implies that $\cdot$ is equal to $\|$.

A Kleene algebra expands a dioid with a star operation satisfying the laws
\begin{align*}
  1 + xx^* &\le x^*, &\qquad z + xy \le y &\implies x^*z \le y,\\
  1 + x^*x &\le x^*, &\qquad z + yx \le y &\implies zx^* \le y.
\end{align*}

For programs, the star represents finite iteration. One can similarly
axiomatise a parallel star operation, but in practice this seems to be
a less useful operation, although one could see it as an operator that
spawns an arbitrary but finite number of processes.

@@

Define a rely-guarantee algebra as a structure
$(K,RG,+,\sqcap,\cdot,\|,^\star,0,1)$, where $(K,+,\sqcap)$ is a
lattice, $(K,+,\cdot,\|,0,1)$ is a weak trioid and
$(K,+,\cdot,^\star,0,1)$ is a Kleene algebra. $RG$ is a distinguished
subset of relys and guarantees which satisfy the following axioms
\begin{align}
r\|r &\le r, \label{rg1}\\
r &\le r\|r', \label{rg2}\\
r\|xy &= (r\|x)(r\|y), \label{rg3}\\
r\|x^+ &\le (r\|x)^+ \label{rg4}.
\end{align}
By convention, we use $r$ and $g$ to refer to elements of $RG$, and
$x,y,z$ for arbitrary elements of $K$. Some elementary consequences of these rules are as follows
\begin{align*}
1 &\le r,\\
r^\star = rr &= r = r\|r,\\
r\|x^+ &= (r\|x)^+.
\end{align*}

@@

Axiom (\ref{rg1}) states that interference from the a rely condition
being run twice in parallel is no different from just the
interference from that rely condition. Axiom (\ref{rg2}) states that
interference from a single rely condition is less than interference
from itself and another rely condition. Axiom (\ref{rg3}) allows a
rely or guarantee to be split across sequential programs. Axiom
(\ref{rg4}) is similar to Axiom (\ref{rg3}) in intent, except it deals
with finite iteration.

@@

Axiom (\ref{rg4}) can be derived from (\ref{rg3}) in all finite
models. This is because any finite $K$ is complete, and in a complete
setting fixpoint laws can be used to show $r\|x^+ \le (r\|x)^+$. Thus
it is impossible to construct a finite model demonstrating that
(\ref{rg4}) is independent from (\ref{rg1}) -- (\ref{rg2}).

@@

Jones quintuples can be encodeded in this setting as
\begin{align}
r, g \vdash \{p\} c \{q\} \iff p(r\|c) \le q \land c \le g.
\end{align}
With this encoding we can derive the standard rely-guarantee inference rules, as
shown in Figure \ref{fig:rgrules}.

@@

\begin{prooftree}
\RightLabel{Star}
\AxiomC{$pr \le p$}
\AxiomC{$r, g \vdash \{p\}x\{p\}$}
\BinaryInfC{$r, g \vdash \{p\}x^\star\{p\}$}
\end{prooftree}
\caption{Rely-guarantee inference rules}
\label{fig:rgrules}
\end{figure}

% Or should this be weak?

While this setting is adequate for deriving the standard inference
rules, the equality is too strong to capture many interesting
statements about concurrent programs. Consider the statement,
$x = y \implies x\|z = y\|z$,
which holds in our algebra and can be read as follows; if $x$ and $y$ are
equal, then they must be equal under all possible interference from
any arbitrary $z$. This would seem to preclude any reasoning about
interference in a fine-grained way, but we can work around this by
using an \emph{interior operator}. In a rely-guarantee algebra an
interior operator is any function $f : K \to K$ which satisfies
\begin{align*}
f(f(x)) &= f(x) \tag{$f$ is idempotent}\\
x \le y \implies f(x) &\le f(y) \tag{$f$ is isotone}\\
f(x) &\le x \tag{$f$ is coextensive}
\end{align*}

We extend our rely-guarantee algebra to
$(K,RG,+,\cdot,\|,^\star,[\--],0,1)$ by adding a single interior
operator $[\--]$, which additionally satisfies
\begin{align}
[xy] &= [x][y], \label{rgimult}\\
[x + y] &= [x] + [y],\\
[x \sqcap y] &= [x] \sqcap [y],\\
[x^\star] &= [x]^\star,\\
[0] &= 0 \label{rgi0}
\end{align}
Intuitively $[x]$ can be thought of as the \emph{execution} of a
program $x$. While $x$ contains all information about what the process
or program \emph{could} do, $[x]$ only contains information about what
it \emph{actually} does. Note that (\ref{rgimult}) -- (\ref{rgi0}) do
not mention parallel composition, such that in general $[x] = [y]
\centernot\implies [x\|z] = [y\|z]$. We amend our jones-quintuples to
involve this operator as shown below.
\begin{align*}
r,g \vdash \{p\} c \{q\} \iff p[r\|c] \le q \land c \le g.
\end{align*}
All the rules in Figure \ref{fig:rgrules} can be derived with this
encoding, and moreover their proofs remain the same, mutatis
mutandis.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algebra for Owicki-Gries}
\label{sec:OG}

This algebra is really unsound!

$px = xp$ is provable!

Following in the style of Kozen's \emph{Kleene algebra with tests}
(KAT), we extend a concurrent Kleene algebra in the same way, yielding
\emph{concurrent Kleene algebra with tests} (CKAT). In this setting we
can encode hoare triples not as $\{x\}y\{z\} \iff xy \le z$, as in
Section \ref{sec:RG}, but instead as
\begin{align*}
\{p\}x\{q\} \iff px \le xq,
\end{align*}
where $p$ and $q$ are KAT-style tests, and not just arbitrary
processes/programs.

Formally, a concurrent Kleene algebra with tests is a structure
\begin{align*}
(K,B,+,\cdot,\|,\overline{\phantom{x}},\phantom{|}^\star,0,1),
\end{align*}
where $(K,+,\cdot,\|,\phantom{|}^\star,0,1)$ is a concurrent Kleene
algebra and $(B,+,\cdot,\overline{\phantom{x}})$ is a Boolean subalgebra such that
\begin{align*}
(p\|q)(x\|y) &= px\|qy,\\ % THESE RULES MAKE NO SENSE!
(x\|y)(p\|q) &= xp\|yq,\\
p\|q &= pq,\\
pp &= p,\\
p &\le 1.
\end{align*}
Just as in KAT, negation is only defined for elements of the boolean
subalgebra.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algebra for Separation Logic}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Language Model}
\label{sec:Model}

In this section, we construct a finite language model satisfying the
axioms in Section \ref{sec:RG}. Restricting our attention to finite
languages means we do not need to concern ourselves with non-algebraic
termination side-conditions, nor do we need to worry about additional
restrictions on parallel composition, e.g. fairness. However, all the
results in this section can easily be adapted to potentially
infinite languages, and our Isabelle/HOL formalisation includes these more
general definitions by using coinductively defined lazy lists to
represent words, and having a weakly-fair shuffle operator for such
infinite languages.

As usual, the product of two languages X and Y over an alphabet
$\Sigma$ is defined as $XY = \{x\frown y.\; x \in X \land y \in Y\}$.
The language product is associative, and distributes over arbitrary
joins from both the left and the right, i.e.
\begin{align*}
X(\bigcup Y) = \bigcup\{XZ.\; Z \in Y\} \qquad \text{and} \qquad (\bigcup Y)X = \bigcup\{ZX.\; Z \in Y\}.
\end{align*}
The shuffle $\sha$ on a pair
of finite words is defined inductively as
\begin{align*}
\epsilon \sha s = \{s\}, \qquad s \sha \epsilon = \{s\},\\
as \sha bt = a(s \sha bt) \cup b(as \sha t),
\end{align*}
which is then lifted to languages X and Y as the shuffle product,
\begin{align*}
X \sha Y = \bigcup\{xs \sha ys.\; x \in X \land x \in Y\}.
\end{align*}
The shuffle product is associative, commutative, and distributes over
arbitrary joins. Both products share the same unit, $\{\epsilon\}$ and
zero, $\emptyset$. Proving properties of shuffle is suprisingly
tricky (especially if one considers infinite words). For a in-depth
treatment of the shuffle product see~\cite{shufflethings}. The
above properties are enough to show that $(\mathcal{P}(\Sigma)^*),
\cup, \cdot,\sha, \emptyset, \{\epsilon\})$ forms a weak trioid.

The rely-guarantee elements in this model are sets containing all the
words which can be built from some set of actions in $\Sigma$. We can
consider a function $\rely : \mathcal{P}(\Sigma) \to
\mathcal{P}(\Sigma)^*$ which takes a set of actions and returns all
the words that can be built from them. This function is an upper
adjoint for $\guar$, the function which returns all the letters in a
language.
\begin{align*}
\guar(X) \le Y \iff X \le \rely(Y)
\end{align*}

Properties (\ref{rg1}) and (\ref{rg2}) from Section \ref{sec:RG} can
mostly be derived from basic properties of $\rely$. Namely,
\begin{align*}
\rely(X \cup Y) &= \rely(X) \| \rely(Y),\\
\qquad \rely(X \cap Y) &= \rely(X) \cap \rely(Y).
\end{align*}
The proofs of (\ref{rg3}) and (\ref{rg4}) are somewhat more involved,
but ultimately rather mechanical, so we do not show them here. We
define the set $RG$ of rely-guarantee elements as the range of the
$\rely$ function and therefore are able to show that $(\mathcal{P}(\Sigma)^*), RG, \cup,
\cdot,\sha, \emptyset, \{\epsilon\})$ is a rely-guarantee algebra.

To make our model more concrete we consider an alphabet comprising
relations ($R,S,T,\dots$) over a state space $\Sigma$. A single
relation represents a single atomic action of a program. Assuming
$\Sigma = \{\sigma_1,\sigma_2,\sigma_3\}$, an example of a word in
$(\Sigma^2)^*$ is shown below.

@@

This word only has a single valid \emph{trace}, which is
$\sigma_1\rightarrow\sigma_1\rightarrow\sigma_2\rightarrow\sigma_3\rightarrow\sigma_2$.
Such state transition traces are commonly called \emph{Aczel
  traces}~\cite{Aczel}. Note that unlike much of the literature involving
Aczel-style traces for concurrency, we omit any kind of process or
environment labelling from our transitions. We define an operator
$\tr(X)$ which returns all such traces in a language $X$. $\tr$ is
both idempotent and isotone.

Note that we cannot use the shuffle operator on directly traces. For
example, imagine we shuffled the single transition $\sigma_2 \to
\sigma_3$ into the above word, we could end up with a word as below.

@@

By performing this shuffle action, we no longer have a trace from
$\sigma_1$ to $\sigma_2$, but instead two traces from $\sigma_2$ to
$\sigma_2$ and $\sigma_3$. This subtle interplay between shuffle and
$\tr$ mimics how Aczel-traces are typically generated from an
operational semantics in much existing rely-guarantee literature. This
operational approach is not desireable from an algebraic perspective
though, as we would rather see programs as mathematical objects in
their own right---a more denotational approach.

Following~\cite{brookes_full_1993} and~\cite{dingel_refinement_2002},
we inductively generate the stutter/mumble language $w^\dagger$ for a
word $w$ in a language over $\Sigma^2$ as follows: Assume $R,S,T \in
\Sigma^2$ and $u,v,w \in (\Sigma^2)^*$. First, $Iw \in
w^\dagger$. Secondly, if $uRv \in x^\dagger$ then $uIRv \in
x^\dagger$ and $uRIv \in x^\dagger$ (\emph{stuttering}). Thirdly, if
$uRSv \in w^\dagger$ then $u(R;S)v \in w^\dagger$
(\emph{mumbling}). The stutter/mumble closure of a language is
thus defined by lifting this operation to languages.
\begin{align*}
X^\dagger = \bigcup\{x^\dagger.\; x \in X\}
\end{align*}

Next, we define the following language for a word $w = R_0R_1\dots
R_n$. The language $w^\udagger$ contains all strings $S_0S_1\dots S_n$
such that for all $m \le n$, $S_m \subseteq R_m$. Effectively this
gives us all the words which are less that $w$ pointwise. Again this
operation is lifted to languages in the same way as before. We define
the set of \emph{$smp$-closed} languages by combining both these
closure operators into a single closure operator, $X^\ddagger$, which
satisfies the properties
\begin{align}
(X^\ddagger \sha Y)^\ddagger &= (X \sha Y)^\ddagger,\\
(X^\ddagger Y)^\ddagger = (&XY)^\ddagger = (XY^\ddagger)^\ddagger,\\
(X \cup Y)^\ddagger &= (X^\ddagger \cup Y^\ddagger),\\
(X^\ddagger \cap Y^\ddagger)^\ddagger &= (X^\ddagger \cap Y^\ddagger),\\
\tr(X^\ddagger) &\subseteq X^\ddagger. \label{trcoext}
\end{align}
These five properties essentially allow us to `lift' all our operators
from above to act on $smp$-closed languages in such a way that they
still form a weak trioid. Henceforth, we will assume all our languages are
$smp$-closed, and all operators have been appropriately lifted.

Equation (\ref{trcoext}) shows that $\tr$ is now an interior operator
on the set of $smp$-closed languages.

We now consider how tests are represented in this model, we define
$\langle R\rangle$ as the singleton language containing just the
relation $R$. Tests are defined as traces of the form $\langle
R\rangle^\ddagger$ where $R \subseteq I$. We denote the set of all
tests as $B$, and let it be ranged over by $P,Q,S$. From the
definition of $smp$-closure $\langle I\rangle^\ddagger =
\{\epsilon\}^\ddagger$, thus the greatest test is $1$, while the least
test is $\langle \emptyset\rangle^\ddagger$, $0$. As such our tests
form a boolean algebra between $1$ and $0$, much they would in
Kozen's Kleene algebra with tests.

\begin{center}
\begin{tikzpicture}[x=1cm,auto]
  \node (center) {};
  \node [elem] (s12) {$\sigma_2$};
  \node [elem, above of=s12] (s11) {$\sigma_1$};
  \node [elem, below of=s12] (s13) {$\sigma_3$};
  \node [right of=s13, node distance=0.75cm] (r1) {};
  \node [below of=r1, node distance=0.5cm] (t4) {$R$};

  \node [elem, right of=s12, node distance=1.5cm] (s22) {$\sigma_2$};
  \node [elem, above of=s22] (s21) {$\sigma_1$};
  \node [elem, below of=s22] (s23) {$\sigma_3$};
  \node [right of=s23, node distance=0.75cm] (r2) {};
  \node [below of=r2, node distance=0.5cm] (t4) {$R$};

  \node [elem, right of=s22, node distance=1.5cm] (s32) {$\sigma_2$};
  \node [elem, above of=s32] (s31) {$\sigma_1$};
  \node [elem, below of=s32] (s33) {$\sigma_3$};

  \node [elem, right of=s32, node distance=1.5cm] (s42) {$\sigma_2$};
  \node [elem, above of=s42] (s41) {$\sigma_1$};
  \node [elem, below of=s42] (s43) {$\sigma_3$};
  \node [right of=s43, node distance=0.75cm] (r4) {};
  \node [below of=r4, node distance=0.5cm] (t4) {$R$};

  \node [elem, right of=s42, node distance=1.5cm] (s52) {$\sigma_2$};
  \node [elem, above of=s52] (s51) {$\sigma_1$};
  \node [elem, below of=s52] (s53) {$\sigma_3$};
  \node [right of=s53, node distance=0.75cm] (r5) {};
  \node [below of=r5, node distance=0.5cm] (t5) {$\lnot R$};

  \node [elem, right of=s52, node distance=1.5cm] (s62) {$\sigma_2$};
  \node [elem, above of=s62] (s61) {$\sigma_1$};
  \node [elem, below of=s62] (s63) {$\sigma_3$};

  \path [line] (s11) -- (s21);
  \path [line] (s12) -- (s22);
  \path [line] (s21) -- (s31);
  \path [line] (s22) -- (s32);

  \path [rel] (s42) -- (s52);
  \path [rel] (s41) -- (s51);
  \path [rel] (s53) -- (s63);
\end{tikzpicture}
\end{center}

A key difference however is that tests only behave as tests when under
$\tr$. This is illustrated above---$\tr(\langle R\rangle^\ddagger\langle
R\rangle^\ddagger) = \tr(\langle R\rangle^\ddagger)$ while $\tr(\langle
R\rangle^\ddagger\langle \lnot R\rangle^\ddagger) = 0$.
